---
title: "Example of loading data and making a masterfile, and protein abundance per MAG"
author: "Magnus Arntzen, Raquel Río"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  pdf_document:
    toc: yes
  html_document:
    theme: united
    toc: yes
    toc_float: yes
---

#Intro

## Load libraries and setup

```{r load libraries, results='hide', message=F, warning=F}
options(stringsAsFactors=F)

library(ggplot2)
library(tibble)
library(tidyverse)
library(ggforce)
library(ggpattern)
library(rlist)
library(gridExtra)
library(pheatmap)
library(patchwork)
library(data.table)
wd <- getwd()

## Load data

```{r load data, results='hide', message=F, warning=F}

read_and_regenerate = F

if (read_and_regenerate) {
  
  #Read DRAM annotations
  #911.040 entries
  genes <- read_delim("DRAM_gene_annotations.tsv", delim = "\t", progress = F, na = c(""))
  colnames(genes)[1] <- "gene"  ##I wrote "gene" because it's the gene associated to that protein
  genes_df <- as.data.frame(genes) %>%
    mutate(Bin = str_split_i(gene, "_contig", 1))
  
  #Read metaproteomics data from FrigPipe
  metaP <- read_delim("Magnus USB/241021_Judith_42696proteins.txt", delim = "\t", progress = F, na = c("", "NaN"), comment = "#")
  metaP_df <- as.data.frame(metaP) %>%
    select(-Description) %>%
    mutate(across(contains("X"), ~ 2^.x)) %>% #we do not want log2 data here
    mutate(Expressed = TRUE) #set all to expressed to ease later filtering of mastertable
  colnames(metaP_df) <- gsub("_2", "", colnames(metaP_df))
  colnames(metaP_df) <- gsub(".Intensity", "", colnames(metaP_df))
  View(metaP_df)
  

  #Make a combined masterfile
  masterfile <- genes_df  %>%
    left_join(metaP_df, by = c("gene" = "Protein")) #All the metaP data is in MetaG but there are metaG data not present in metaP, but we want to keep all
  View(masterfile)  
  
  # This masterfile has:
  #    911,040 entries, 60 samples: 29 stress, 31 control.
  #    42,696 expressed proteins (at least one value)

  write_delim(masterfile, file = "masterfile_genes_proteins_gutbrain.txt", delim = "\t")
  saveRDS(masterfile, file="masterfile.rds")

} else {
  masterfile <- readRDS(file="masterfile.rds")
}


#Here it is possible to also read in the taxonomy og the MAGs fr use later


```

## Protein abundance per bin

```{r protein abundance, fig.height = 60, fig.width = 4, results='hide', message=F, warning=F}

#this will generate a list of proteins that are expressed,
#removing proteins that are indistinguishable to proteins from other MAGs
#but keeping them if they are from the same MAG
proteins <- masterfile %>%
  filter(Expressed == TRUE) %>%
  select(-ko_id, -kegg_hit, -pfam_hits, -cazy_ids, -cazy_best_hit, -Expressed) %>%
  rowwise() %>%
  mutate(Ind_bins_temp = str_split(`Indistinguishable Proteins`, ", ")) %>%
  mutate(Indistinguishable_bins = paste0(unique(str_split_i(Ind_bins_temp, "_contig", 1)), collapse=";")) %>%
  mutate(Indistinguishable_bins = na_if(Indistinguishable_bins, "NA")) %>%
  ungroup() %>%
  select(-Ind_bins_temp) %>%
  filter(is.na(`Indistinguishable Proteins`) | Bin == Indistinguishable_bins) 

View(proteins)

#This is 39.406 proteins from 481 bins (n_distinct(proteins$fasta))

protein_sums_per_bin <- proteins %>%
  select(-`Indistinguishable Proteins`, -Indistinguishable_bins, -fasta) %>% ###I remove fasta because it's the same as Bin
  pivot_longer(cols = 12:71) %>%  ######I selected the "protein abundance for each sample" columns
  group_by(Bin, name) %>%
  summarise(summedLFQ = log2(sum(value, na.rm = T))) %>%
  mutate(summedLFQ = if_else(summedLFQ == "-Inf", 0, summedLFQ)) %>%
  pivot_wider(values_from = summedLFQ)

View(protein_sums_per_bin)
write_delim(protein_sums_per_bin, file = "summedLFQ_per_bin.txt", delim = "\t")  

pheatmap(protein_sums_per_bin %>% column_to_rownames("Bin"), fontsize = 4)




```

